Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-25 manuscript/manuscript.Rmd
datafreeze: modified code/*.py files to create data in a standardized form. created final data files data/*.xz, removed previous data/*.gz files (still probably need to remove them from history with git filter-branch (TBD). Modified code blocks within manuscript.rmd to fit the new data form, things are stable here.

```{r}
# Collect data (outsource to separate R script later TBD)
msy <- read_csv("../data/msy.csv.gz")
ppo <- read_csv("../data/PPO200.csv.gz")
escapement <- read_csv("../data/escapement.csv.gz", show_col_types = FALSE)

# rename some columns for figure purposes:
colnames(msy) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z')
colnames(ppo) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z') 
colnames(escapement) <- c('t', 'rep', 'esc_level', 'reward', 'X', 'Y', 'Z') 
opt_escapement <- escapement |> filter(esc_level == 0.62)

# ppo dataset uses step rewards rather than cumulative rewards, which is what is useful here.
# so, switch stuff around:
#ppo |> 
#    group_by(rep) |> 
#    mutate(cumulative_reward = cumsum(reward)) |>
#    within(rm("reward", "1")) |>
#    rename(reward = cumulative_reward)

@@

# best constant effort based on the data on ../data/msy.csv.gz
best_action <- 0.045
msy_sim <- msy |>
    filter(action == best_action) |> 
    group_by(rep) |>
    pivot_longer(c("X", "Y", "Z"),
                names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "msy") # natural units, policy name

actions <- unique(msy$action)
i <- which.min(abs(actions - best_action* 0.8))
tac_sim <- msy |> 
    filter(action == actions[[i]]) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "tac")

ppo_sim <- ppo |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "ppo")

best_e <- 0.61 # optimal constant escapement, computed from ../data/escapement.csv.gz
esc_sim <- escapement |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1)

@@

```{r}
msy_rewards <- msy |>
  group_by(action, rep) |>
  filter(t == max(t)) |> 
  group_by(action) |>
  summarise(mean_reward = mean(reward),
            sd = sd(reward))

esc_rewards <- escapement |>
  group_by(esc_level, rep) |>
  filter(t == max(t)) |> 
  group_by(esc_level) |>
  summarise(mean_reward = mean(reward),
            sd = sd(reward))

@@

```{r timeseries, fig.cap="TBD.", out.width = "6.5in"}
policy_labels <- list("msy" = "Opt. Const. Mortality", "tac"= "80% Opt. Const. Mortality", "ppo"= "DRL")
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 5) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/3)
```

@@

```{r timeseries_old, fig.cap="TBD.", out.width = "6.5in"}
policy_labels <- list("msy" = "Opt. Const. Mortality", "tac"= "80% Opt. Const. Mortality", "ppo"= "DRL")
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 5) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/3)
```

@@

pesc <- opt_escapement |>
  mutate(X = X+1, Y = Y+1, Z = Z+1, 
        harvest = case_when(
          X - esc_level > 0 ~ X - esc_level,
          X - esc_level <= 0 ~  0)
         ) |> 
  ggplot(aes(X, Z, color = harvest)) +
  geom_point() + sc +
  ggtitle("Opt. Const. Escapement")

@@

ppo_mod <- ppo |> 
  mutate(X = X+1, Y = Y+1, Z = Z+1, escapement = (1-action)*X, time = t)

>>>>

```{r}
# for best const mortality and best const escapement
msy_optimizer <- function(df) {
  df |> 
  group_by(action, rep) |>
  filter(t == max(t)) |> 
  group_by(action) |>
  summarise(mean_reward = mean(reward), sd = sd(reward)) |> 
  filter(mean_reward == max(mean_reward))
}
escapement_optimizer <- function(df) {
  df |> 
  group_by(escapement, rep) |>
  filter(t == max(t)) |> 
  group_by(escapement) |>
  summarise(mean_reward = mean(reward), sd = sd(reward)) |> 
  filter(mean_reward == max(mean_reward))
}

# Collect data (outsource to separate R script later TBD)
msy_df <- read_csv("../data/msy.csv.xz")
best_action <- msy_optimizer(msy_df)$action
escapement_df <- read_csv("../data/escapement.csv.xz", show_col_types = FALSE)
best_e <- escapement_optimizer(escapement_df)$escapement
ppo_df <- read_csv("../data/PPO200.csv.xz")

opt_escapement <- escapement_df |> filter(escapement == best_e)

@@

msy_sim <- msy_df |>
    filter(action == best_action) |> 
    group_by(rep) |>
    pivot_longer(c("X", "Y", "Z"),
                names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "msy") # natural units, policy name

actions <- unique(msy$action)
i <- which.min(abs(actions - best_action* 0.8))
tac_sim <- msy_df |> 
    filter(action == actions[[i]]) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "tac")

ppo_sim <- ppo_df |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "ppo")

best_e <- 0.61 # optimal constant escapement, computed from ../data/escapement.csv.gz
esc_sim_df <- escapement |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1)

@@

```{r}
msy_rewards <- msy_df |>
  group_by(action, rep) |>
  filter(t == max(t)) |> 
  group_by(action) |>
  summarise(mean_reward = mean(reward),
            sd = sd(reward))

esc_rewards <- escapement_df |>
  group_by(escapement, rep) |>
  filter(t == max(t)) |> 
  group_by(escapement) |>
  summarise(mean_reward = mean(reward),
            sd = sd(reward))

@@

```{r timeseries, fig.cap="TBD.", out.width = "6.5in"}
policy_labels <- list("msy" = "MSY Const. Mortality", "tac"= "80% MSY Const. Mortality", "ppo"= "DRL")
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 5) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/3)
```

@@

```{r timeseries_old, fig.cap="TBD.", out.width = "6.5in"}
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 5) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/3)
```

@@

pesc <- opt_escapement |>
  mutate(X = X+1, Y = Y+1, Z = Z+1, 
        harvest = case_when(
          X - escapement > 0 ~ X - escapement,
          X - escapement <= 0 ~  0)
         ) |> 
  ggplot(aes(X, Z, color = harvest)) +
  geom_point() + sc +
  ggtitle("Opt. Const. Escapement")

@@

ppo_mod <- ppo_df |> 
  mutate(X = X+1, Y = Y+1, Z = Z+1, escapement = (1-action)*X, time = t)
